\section{Conclusion}

Adding a finite speed of information to the self-replicating slingshot model turns a convenient
assumption into a measurable cost, but not the cost a coarse model reports. Once the dynamics are
resolved so that each probe decides at its true arrival time, the tens-of-percent fill-time penalty
seen in a fixed-timestep model collapses -- it was a discretization artifact -- and the slingshot
speed-up survives intact, reproducing Nicholson and Forgan quantitatively. The genuine cost is
redundant travel: every star is still settled, with a negligible change in the number of probes
built, but stale views send probes to stars others have already claimed. That waste follows a
simple law we both derive and measure -- the delay multiplies the wasted journeys by $1 + v/c$, so the
tax, the extra wasted trips as a fraction of those a perfect-information swarm already flies, equals
the probe's speed in units of light, $\Lambda = v/c$ -- from negligible at the powered cruise, through
about a percent at slingshot speeds, to a fifth at the directed-energy speeds beamed propulsion
targets. (The baseline is itself wasteful, about four in five journeys even with a perfect map, so
this is a $v/c$ surcharge, not a fifth of all trips.) The law holds under a clumpy field, where if
anything it softens; it grows with the replication branching factor without saturating up to sixteen
offspring; and whether it also costs energy is mission-dependent: only a rendezvous that must brake
pays for a wasted trip's kinetic energy, which for the fast policies was largely gravitationally
sourced rather than fuelled. The cost is dominated
by fuel and effort; a secondary fill-time penalty of a few percent rides alongside it, but only in
the fast directed-energy regime and never at slingshot speeds. It is not a hard floor: a probe that
listens in flight, one of the two extensions the model invites, recovers essentially all of the
wasted-arrival tax at a small cost in fill time, while a full probe-to-probe relay of the settled
map remains open. Neither can remove the physical floor beneath the tax, and how much either can buy
back is itself bounded by the finite speed of light.

\section*{Data Availability}
The model is a deterministic, seeded, event-driven fold; every figure, table, and reported
statistic regenerates from source by running the swarm module's experiment and figure scripts,
and every quantitative claim traces to a cited source in the project's shared bibliography. The
code and its citation metadata are archived with the concept DOI \texttt{10.5281/zenodo.21249296}.

\section*{Acknowledgment}
This work used no external funding.
