\section{Discussion}

The result refines rather than overturns the slingshot picture, and it does so twice over.
Nicholson and Forgan's speed-up is real, large, and here reproduced to the same order of magnitude
\citep{nicholson-forgan-2013}; finite light-speed does not erode it, and does not measurably
delay the fill of the field. What a perfect-information model hides is not lost time but lost
motion: probes flying to stars that others have already claimed. That redundant travel is the
coordination tax, it is governed by the probe's speed in units of $c$, and it matters only once
a swarm flies fast enough for a decision to outrun its own news.

\subsection{What the tax is denominated in}
Because the field always fills and each settlement launches a fixed number of offspring, the two
coordination regimes build essentially the same number of probes, differing by well under a
percent; the delay changes almost nothing about how much is manufactured, only how far it is
flown. The tax is therefore a travel cost, and we deliberately denominate it in journeys rather
than energy, because translating journeys into energy is mission-dependent in a way our model does
not fix. A wasted journey's kinetic energy is not straightforwardly a fuel cost: under the
slingshot policies most of that energy was extracted gravitationally from stellar motion rather
than spent as propellant, and a probe that re-targets on arrival keeps its speed and pays nothing
further. The genuine energy penalty of a wasted trip is whatever a rendezvous at the doomed star
would have cost -- braking to rest and re-accelerating -- which scales with $v^2$ for a fuel-braked
craft but is near zero for a flyby or a gravitationally-braked one. We therefore do not headline an
energy figure; the honest statement is that the journey tax becomes an energy tax only to the
extent a mission must brake, and then it grows with $v^2$. A caveat cuts the other way, too: even as
a travel cost the tax is denominated in journeys, not in the swarm's total resource budget, which
is dominated by building the probes -- a cost we hold fixed and identical across regimes and do not
model. As a fraction of total mission resources the coordination tax is smaller than its share of
journeys.

\subsection{Limitations}
\label{sec:limitations}
The decision-site headline is a conservative upper bound: probes here learn only at their decision
stars, ignoring news their paths cross in flight. We relax exactly that assumption in
Section~\ref{sec:floor}, where a probe that listens in flight recovers essentially all of the
wasted-arrival tax, so the headline is an upper bound rather than a floor. A true probe-to-probe
relay of the settled map, which we do not implement, would sit between the beacon-only in-flight
bound and the decision-site one and is the natural next step. The floor beneath all of them -- that
no probe can know a distant star is taken until its light arrives -- is set by physics, not
protocol.

A further, inherited idealization concerns the stars themselves. They carry the velocity vectors
that power each slingshot but are held fixed in position, exactly as in the two source models,
both of which name this as their single most important simplification
\citep{nicholson-forgan-2013,forgan-papadog-kitching-2013}. It is not a small one: at the local
circular speed a star sweeps about $225$ pc in a million years, and even the peculiar velocity
dispersion alone carries it some $30$ to $40$ pc \citep{kerr-lynden-bell-1986,nordstrom-2004-gcs},
far past the roughly one-parsec mean hop, so over a Myr-scale fill real proper motion would
reshuffle which stars are near which. One might object that this touches the delayed-belief gate
itself, not just the backdrop: the beacon leaves where the star \emph{was}, so a moving field would
need retarded positions in (\ref{eq:gate}). But that specific correction is tiny, and for a clean
reason: during a beacon's light-crossing a star moves only the fraction $v_\star/c$ of the distance
it is signalling across, where $v_\star$ is the \emph{relative} velocity between emitter and
observer -- the bulk Galactic rotation is common to both and cancels, leaving the peculiar and shear
velocity, $\sim\!40~\mathrm{km/s}$, so $v_\star/c \approx 10^{-4}$, a hundredth of a percent whatever
the decision distance (and even the full circular speed, $220~\mathrm{km/s}$, gives $7\times10^{-4}$,
under a tenth of a percent) -- so the gate reads essentially the right geometry at the moment of
decision. What the frozen field removes is the slow, Myr-scale drift
of the layout, and that is exactly what the paired design controls: both coordination regimes run on
the \emph{identical} frozen field, so the difference still isolates the light-lag cost even where the
absolute layout is idealized. This cancellation extends to the \emph{dynamics}, not just the static
geometry: what the tax multiplies is the baseline per-hop contention $p_\text{perfect}$, set by the
local stellar density and the branching factor, both of which are mode-independent and both of which
see the identical drift under either regime -- so a moving field would perturb the paired difference
only at second order, through any correlation between a star's motion and its being contended, which
the global constancy of $v$ and $c$ already suppresses in the same way it suppresses the density
dependence (Section~\ref{sec:clumpy}). A direct moving-field spot-check would confirm this and is the
natural next test; the absolute fill times, meanwhile, do inherit the idealization. The
slingshot is idealized in a second way we inherit from the source models: we track the scalar speed a
probe gains from an encounter and let it then depart toward its chosen target, rather than resolving
the outgoing \emph{direction} that a real gravitational assist fixes by geometry. A directional model
would couple boost to heading and so lower the achievable boosted speed, and thus $\Lambda$, in the
slingshot regime; but because that regime already pays only about a percent, and the headline $v/c$
law is driven by the directly-swept powered speed rather than by slingshot mechanics, the
simplification does not touch the law.

One idealization we do \emph{not} need to correct is relativity, and for a stronger reason than
smallness. The tax counts journeys, and every quantity that enters it -- the belief gate
(\ref{eq:gate}), the travel time $d/v$, the light-delay $d/c$, and the in-flight learn time of
Section~\ref{sec:floor} -- is a coordinate time in the one shared galaxy frame, the frame in which
the star field is at rest and the beacons are defined. The governing ratio $\Lambda = (d/c)/(d/v) =
v/c$ is a ratio of two coordinate times in that single frame; no Lorentz transformation to the probe
frame enters, and $v < c$ throughout, so $\Lambda = v/c$ is \emph{exact} for the coordination tax at
any sub-luminal speed, not a small-$\beta$ linearization. The Lorentz factor at $0.2\,c$ ($\gamma
\approx 1.02$, a kinematic correction of about $2\%$) touches only the separate energy weighting,
where it never enters a headline number, and not the journey count.

Three further bounds on scope. First, the field is uniform by default; changing its density only
rescales the clock (confirmed), and a clumpy, non-uniform field leaves the law intact -- the
redundant-travel tax stays $\approx \Lambda$, softening if anything at extreme substructure
(Section~\ref{sec:clumpy}). What we do not probe is a full galactic disk with a large-scale radial
density gradient. Second, the system-size sweep now spans about $1024\times$, from 256 to 262{,}144
stars: serving the nearest-not-yet-settled query with a dynamic spatial index over the unsettled set
makes a single run near-linear in $N$, lifting the quadratic cost that once held the sweep to a
sixteen-fold range. Over this longer arm the tax does not grow with scale -- it holds near a fifth up
to about a thousand stars and then falls steadily to about one part in eighty at 262{,}144, a
genuine bulk decline that survives both an interior-only masking and a periodic-box control and so is
not a hard-wall boundary artifact (Section~\ref{sec:scale}) -- so we still make no quantitative
extrapolation to galactic ($\sim\!10^{11}$-star) scale. A $1024\times$ lever arm is far longer than a
sixteen-fold one but remains five orders of magnitude short of a real galaxy, and in any case a range
over which the fraction only shrinks cannot support carrying a fixed percentage outward. Third, the
dynamics have no loss term: settled sites never fail, so the field always closes. A
settlement-death term of the kind that produces the ``Aurora'' steady state
\citep{carroll-nellenback-2019} is what would be needed to turn coordination failure into a
permanent unsettled remainder; delay alone does not.

\subsection{Relation to delayed agreement}
This is a coordination problem, and the distributed-systems literature names the regime it sits in --
by analogy, not by derivation; we invoke it only to place that regime, and the result itself is our
own. When communication delay grows past the decision cadence a system can no longer maintain a single
agreed state and must trade consistency for availability, acting on local views
\citep{gilbert-lynch-2002,olfati-saber-fax-murray-2007}. At parsec spacing the round-trip light-time
is years, far longer than any plausible decision cadence, so a swarm that must keep acting has no
option but availability -- which is exactly what the gate (\ref{eq:gate}) encodes, each probe an
independent colony acting on priors. We borrow only this picture. The classical impossibility results
assume fault models that are not ours (asynchronous crash faults for FLP \citep{flp-1985}, partition
for CAP \citep{gilbert-lynch-2002}) and none of them predicts the $\Lambda = v/c$ scaling, which is at
bottom a geometric-probabilistic result about stale target selection, not a consensus theorem. The
mechanism itself -- agents on a stale shared map flying redundant trips to targets others have already
claimed, and communication cutting that overlap -- is familiar from multi-robot exploration, where
coordinated frontier assignment exists precisely to suppress it \citep{burgard-2005,amigoni-2017}.
What is specific here is not that stale views waste trips but that, when the delay is light-speed
itself, the wasted fraction takes the clean form $\Lambda = v/c$, independent of hop length and
density. The coordination tax is simply what that enforced independence costs when probes want the
same stars.

\subsection{An implication for design}
The corollary is a threshold, not a preference. At slingshot speeds the tax is about a percent, so
coordination machinery -- relay, gossip, a shared map -- returns less than its mass costs, and the
coordination overhead is negligible against what speed buys. At the directed-energy speeds that make
a probe fast enough to matter, the same physics that buys the speed also makes stale views expensive:
a fifth more journeys are wasted, of which -- as Section~\ref{sec:floor} shows -- a probe that merely
listens in flight recovers nearly all. The implication is that the value of coordination scales with
how fast the swarm flies, while the finite speed of light sets only the small residual -- the travel
already flown before the news arrives -- that no protocol can ever recover.
