\section{Introduction}

A self-replicating probe does not explore a galaxy alone. It settles a
star, builds copies of itself from local material, and sends them on to further stars,
so a single launch grows into a front that sweeps outward and fills the reachable field.
That a machine can build a copy of itself at all is von Neumann's result
\citep{von-neumann-burks-1966}, and a probe of
this kind saturating the Galaxy in a few million years is the sharpest form of the Fermi
question: if interstellar settlement is possible it should already be complete, so the
observed silence is a paradox \citep{hart-1975,tipler-1980}. How long the filling takes is
the exploration-timescale question, and it has a long modelling lineage. Early work framed
each interstellar probe as an independent agent acting on priors fixed before launch
\citep{freitas-1980-interstellar-probe}, and treated settlement as a front: a
density-dependent diffusion with a travelling-wave solution \citep{newman-sagan-1981}, or a
Monte Carlo migration filling the Galaxy in tens of millions of years \citep{jones-1981}.
Later simulations made the field concrete, sending probes and subprobes through a defined
stellar volume to measure fill times directly \citep{bjork-2007,cotta-morales-2009}, and
established the seeded Monte Carlo methodology that situates an exploration timescale
within a distribution of realizations rather than a single number \citep{forgan-2009}.

The sharpest treatment for self-replicators is the slingshot line.
\citet{forgan-papadog-kitching-2013} showed that a single probe stealing energy from the motion of
the stars it passes, a gravitational slingshot, explores up to two orders of magnitude
faster than a probe under its own power. \citet{nicholson-forgan-2013} extended that manoeuvre to
self-replicating probes and confirmed the speed-up survives replication, with the surprising result
that aiming for the nearest star can beat aiming for the largest boost. The probe speeds that make this
regime interesting, from the powered cruise of $3\times10^{-5}\,c$ up toward the $0.1$ to
$0.2\,c$ of beamed directed-energy concepts \citep{lubin-2016}, are exactly where the next
assumption starts to hurt.

That assumption, which Nicholson and Forgan name as a limitation and leave for future
work, is that every probe is granted perfect, instantaneous knowledge of which stars are
already settled. No probe ever flies to a star another has just claimed, because it always
knows. In a real galaxy no such oracle exists: news that a star has been settled can
travel no faster than light, so a probe choosing its next target acts on a picture of the
field that is, for distant stars, years or centuries out of date. This is the elementary
consequence of a finite signal speed, that events have only a partial ordering and a
remote event is unknowable until its signal arrives \citep{lamport-1978}.

This paper adds that light-speed limit to the self-replicating slingshot model and measures
what it costs. The answer, built from a controlled paired-ensemble experiment, has two parts,
and the first is a caution. The cost is easy to overstate: a simulation that advances in a
fixed timestep coarser than the interval between probe launches makes many probes decide from
the same stale snapshot at once, and they collide far more than they physically would. Read
that way the delay appears to slow the fill of the field by tens of percent. That number is an
artifact. When the dynamics are resolved so that each probe decides at its own arrival time, the
tens-of-percent slowdown collapses to a few percent at most, and to nothing at slingshot speeds,
while the slingshot speed-up survives contact with finite information intact, reproducing the
source model to the same order of magnitude. The dominant cost shows up elsewhere. Every star still gets settled,
and with a negligible change in how many probes are built, well under a percent; what the delay
chiefly changes is how many wasted journeys the swarm flies to stars other probes have already
claimed. That is a cost paid mostly in fuel and effort rather than time, and a
perfect-information model hides it entirely. It is governed cleanly by a single dimensionless
number, the probe's speed in units of the speed of light. The delay multiplies the wasted
journeys by $1+\Lambda$, so the tax -- the extra wasted trips as a fraction of those a
perfect-information swarm already flies -- is negligible for a powered cruise, about a percent at
slingshot speeds, and a fifth at the directed-energy speeds that beamed-propulsion concepts target.
That baseline is itself wasteful: even with a perfect map about four in five journeys are
redundant, so the tax is a $v/c$ surcharge on an already-redundant process, not a fifth of all
journeys. We derive from a short collision argument that this relative tax should equal $\Lambda$
itself and confirm it in simulation, show the tax grows with the replication branching factor
without saturating up to sixteen offspring and survives a clumpy (non-uniform) field, and -- because a probe deciding only
at its stops is pessimistic -- that a probe which listens in flight recovers almost all of it,
bounding how much coordination can ever buy back. The cost is dominated
by fuel and effort, with a secondary fill-time penalty that emerges only at the highest speeds; and
it scales, cleanly, with how fast the swarm flies.
