\section{The light-delayed belief model}
\label{sec:model}

\subsection{The knowledge gate}
The change to the model is one gate on knowledge. A settled star is treated as an
omnidirectional beacon that begins emitting in the year it is settled. A probe deciding its
next target at star $f$ in year $Y$ believes a distant star $i$ is settled only once that
beacon's light has had time to arrive:
\begin{equation}
\text{settled}(i) + \frac{\mathrm{dist}(f, i)}{c} \le Y . \label{eq:gate}
\end{equation}
Under the perfect-information assumption of the source model, (\ref{eq:gate}) collapses to
``settled at any non-negative year,'' and the model reproduces our perfect-information
baseline bit-for-bit; the light-speed term is what introduces stale views. When two probes
both believe an unsettled star is the best target they both fly to it; the loser arrives to
find it taken, its trip wasted, and must re-target from the even more stale picture it holds on
arrival. This is availability chosen over consistency: rather than wait for a single agreed map
that finite light-speed makes impossible, each probe acts now on its local view
\citep{lamport-1978}. Stars carry velocity magnitudes that feed the slingshot boost but are held
fixed in position, following the source model, so $\mathrm{dist}(f,i)$ in (\ref{eq:gate}) is
static and the beacon geometry does not shift under the probe. Physical arrival always reads
ground truth: what a probe finds on arrival is the star's real status, whatever it believed
when it aimed.

\subsection{The three flight policies, and the speed axis}
We carry three named policies, following the slingshot studies we extend
\citep{forgan-papadog-kitching-2013,nicholson-forgan-2013}. \emph{Powered nearest-neighbour}
flight moves under the probe's own power to the nearest unsettled star. \emph{Nearest-star
slingshot} flight takes the same nearest target but gains speed from gravitational slingshots
off the stars it passes. \emph{Maximum-boost slingshot} flight instead selects, among the
nearest believed-unsettled stars, the target that yields the largest velocity gain. The
gravitational boost self-limits (its per-encounter gain falls off once the probe outruns the
stellar escape speed), so both slingshot policies saturate at a mean speed of a few thousand
km/s, about $10^{-2}$ of $c$. To reach the higher speeds that beamed directed-energy
concepts target, $0.1$ to $0.2\,c$ \citep{lubin-2016}, a probe must be
driven, not slung; we therefore sweep the powered cruise speed as a clean, direct control on
$v/c$, from the source model's $3\times10^{-5}\,c$ up to $0.2\,c$. Powered flight thus spans the
whole speed axis, while the two slingshot policies sit near $v/c \approx 0.01$.

\subsection{Event-driven resolution, and why it is necessary}
The fold can advance either by a fixed timestep or by jumping to the next probe arrival. This
is not a mere numerical convenience, and getting it wrong is the single largest source of error
in this problem. A fixed timestep $\Delta t$ processes together every probe that arrives within
a window, so all of them decide against the same snapshot of the settled map. When $\Delta t$
exceeds the interval between launches -- which it does badly in the boosted regime, where a
probe crosses a parsec in centuries against a source-model step of thousands of years -- the
window batches many independent decisions into one synchronized burst, and they collide far
more often than they would if each decided at its true arrival time. The measured coordination
cost then reflects the size of the timestep, not the physics. We therefore run the model
event-driven throughout: the fold jumps to the earliest outstanding arrival, advances the clock
to it, and processes exactly the probes arriving then. This is the exact $\Delta t \to 0$ limit,
dt-independent by construction, and it is what every number in this paper is computed on.
Section~\ref{sec:results} shows directly that the apparent fill-time tax is an artifact of a
coarse step and collapses to zero as the step resolves.

\subsection{The governing ratio}
How much the gate matters is governed by one dimensionless number: the information delay across
a hop divided by the travel time across it. For a hop of length $d$ the delay is $d/c$ and the
travel time is $d/v$, so the ratio is
\begin{equation}
\Lambda \;=\; \frac{d/c}{d/v} \;=\; \frac{v}{c}, \label{eq:lambda}
\end{equation}
the probe's speed in units of $c$, independent of the hop length. A slow probe
($\Lambda \ll 1$) outpaces its own ignorance only trivially, so stale views barely arise; a
fast probe carries a large $\Lambda$ and routinely decides before the relevant news can reach
it. At the powered cruise speed $\Lambda \approx 3\times10^{-5}$ is negligible.

\subsection{A collision model: why the tax should equal $\Lambda$}
\label{sec:theory}
The same ratio predicts the size of the redundant-travel tax, not just its trend, from a short
exposure argument. Fix a probe that has committed to a target star at distance $d$. Under perfect
information it chose a truly unsettled star, so it is wasted only if some other probe claims that
star during its flight -- an exposure window equal to the travel time $d/v$. Under the light-speed
gate the probe carries an \emph{additional} blindness: the target may already have been claimed up
to a light-time $d/c$ before launch, with the beacon not yet arrived, and the probe would have
launched anyway. Its total exposure to an unseen claim is therefore $d/v + d/c$. If claims of a
given candidate arrive at some roughly steady local rate, the probability of a wasted trip is
proportional to the exposure window, so the ratio of wasted trips with and without the gate is
\begin{equation}
\frac{p_{\text{lag}}}{p_{\text{perfect}}} \;\approx\; \frac{d/v + d/c}{d/v} \;=\; 1 + \frac{v}{c}
\;=\; 1 + \Lambda,
\label{eq:tax}
\end{equation}
and the coordination tax -- the extra wasted trips as a fraction of the perfect-information
baseline -- is just $p_{\text{lag}}/p_{\text{perfect}} - 1 \approx \Lambda$. Two features of
(\ref{eq:tax}) are worth stating because the simulation independently reproduces both: the hop
length $d$ cancels, so the tax cannot depend on how far apart the stars are, and the local claim
rate cancels, so it cannot depend on the stellar density. This is a leading-order estimate -- it
assumes claims of a candidate are independent and locally uniform in time, and it ignores both the
geometry of which stars are in contention and the saturation of already-doomed hops -- so it predicts
the coefficient of proportionality is one, as a ceiling that those corrections can only lower.
Section~\ref{sec:results} shows the measured tax follows $\Lambda$ in \emph{form} but with a
coefficient below that ceiling: a through-origin slope of $0.82$ on the $N = 2{,}048$ headline field,
resolvably under one and declining with system size as saturation strengthens (from $a \approx 0.97$
at a few hundred stars to $0.068$ at a quarter-million). The simulation confirms the $d$- and
density-independence of the form directly.

\subsection{Experimental design and parameters}
The field is a box of stars at a uniform density of one star per cubic parsec, the convention of
the source model. Each settled star launches offspring (the replication branching factor):
branching is what makes two probes ever contend for a star, so it is the precondition for any
coordination effect at all and the parameter most likely to change the answer, so we take two
offspring as the default and sweep it explicitly (Section~\ref{sec:branching}). Three properties
keep the measurement honest. First, knowledge is evaluated only at the decision star at the moment
of decision; news a probe's path happens to cross in mid-flight is ignored, which deliberately
undercounts what a probe could know and so makes the measured waste a conservative upper bound --
we measure the opposite, optimistic bound (a probe that listens in flight) separately in
Section~\ref{sec:floor}. Second, a probe that loses a race re-targets, but only up to a fixed
number of times before it is retired; this cap is applied identically under both coordination
regimes, so the paired comparison is fair. On the few-hundred-star fields of earlier work the tax
plateaued above about $\text{cap} = 8$, which is why we adopt $\text{cap} = 8$ as the default -- but
that plateau does not survive to larger fields (Section~\ref{sec:scale-companions}): at $\Lambda =
0.2$ and $N = 262{,}144$ the fuel tax is $0.27$, $0.27$, $1.31$, $4.79$, $8.97\%$ for caps of
$2, 4, 8, 16, 32$, still climbing with no sign of levelling. The default cap of $8$ is therefore a
documented \emph{lower bound}, not a converged value, and every fuel-tax figure it produces is an
underestimate whose size grows with the field (negligible at small $N$, substantial at scale). Third, the whole simulation is a pure,
seeded, event-driven fold: fixing the seed fixes the star field, every arrival, and every target
choice, so a run with the light-speed gate and a run without it can be compared on the
\emph{identical} galaxy. That paired design attributes a difference to the delay alone rather than
to the luck of the star layout.

Two dependencies are worth stating because they are easy to over-read. Changing the value of the
uniform density rescales every inter-star distance by a common factor, multiplying travel time and
light-travel time on every hop equally, so $\Lambda$ and the relative penalties are unchanged: this
is a scale-free property of a uniform field, exactly what (\ref{eq:tax}) predicts when the hop
length cancels, so confirming it holds to the printed digit from $0.14$ to $5$ stars per cubic
parsec \citep{recons-census} is a consistency check on the derivation, not an independent robustness
result. The case that would genuinely test density sensitivity is a clumpy, non-uniform field, which
we examine directly in Section~\ref{sec:clumpy}; what we do not probe is a full galactic disk with a
large-scale radial density gradient. And because the field always fills under
both regimes and each settlement launches a fixed number of offspring, the two regimes build
essentially the same number of probes -- they differ only in a handful of terminal launches as
targets run out, well under a percent -- so the resource the delay actually moves is the number
of journeys flown, which is why the coordination cost is a travel cost.

The ensemble size varies by measurement, sized to each one's event-mode cost; Table~\ref{tab:ensembles}
lists them. Every run is event-driven and paired -- each seed fills the identical field under both
coordination regimes -- and the seed counts are prefixes of one fixed 64-seed set, so the sweeps share
seeds and differ only in the stated star count.

\begin{table}[!t]
  \centering
  \caption{The paired ensemble used by each measurement. All are event-driven and paired (each seed
    filled under both coordination regimes on the identical field); seed counts are prefixes of one
    fixed 512-seed set, sized to each measurement's cost. Star counts are powers of two computed by
    the \texttt{run\_fill\_flat} fast path (Issue \#38); each fixed-$N$ sweep runs on the largest
    power-of-two field it can complete in about a minute of wall-clock, so the measurements sit at
    different $N$ rather than a shared anchor (two exceptions: the clumpy-field sweep is kept on its
    historical $N = 500$ ensemble, its scale companion having no power-of-two rerun; and the
    re-target-cap sweep runs at the top of the ladder, $N = 262{,}144$, because its whole purpose is
    to test whether the cap plateau survives to scale). The lower block
    repeats each fixed-$N$ sweep at $N = 262{,}144$ (the clumpy-field companion excepted -- it stays
    at $N = 200{,}000$) using the finite-size arm's precision-target seed
    count, so every claim rests on a scale companion in addition to its headline result -- discussed
    together in Section~\ref{sec:scale-companions}.}
  \label{tab:ensembles}
  \begin{tabular}{lccl}
    \toprule
    Measurement & Seeds & Stars & Reported in \\
    \midrule
    Speed sweep (headline)  & 512 & 2{,}048 & Table~\ref{tab:lambda}, Fig.~\ref{fig:lambda} \\
    Branching sweep         & 32 & 8{,}192 & Fig.~\ref{fig:branching} \\
    Re-target-cap sweep     & 32 & 262{,}144 & Section~\ref{sec:results} \\
    In-flight floor         & 48 & 32{,}768 & Fig.~\ref{fig:floor} \\
    Concurrency             & 16 & 32{,}768 & Fig.~\ref{fig:concurrency} \\
    Finite-size sweep       & 48 down to 8 & 256--262{,}144 & Fig.~\ref{fig:scale} \\
    Finite-size, interior edge test & 32 down to 6 & 256--262{,}144 & Section~\ref{sec:scale} \\
    Finite-size, periodic control   & 32 down to 6 & 256--262{,}144 & Section~\ref{sec:scale} \\
    Clumpy-field sweep      & 48 & 500 & Fig.~\ref{fig:clumpiness} \\
    Timestep artifact       & 32 & 8{,}192 & Fig.~\ref{fig:dt} \\
    \midrule
    \multicolumn{4}{l}{\emph{Scale companions, precision-target seed count (Section~\ref{sec:scale-companions})}} \\
    Speed sweep, at scale   & 8  & 262{,}144 & Section~\ref{sec:scale-companions} \\
    Re-target-cap, at scale & 8  & 262{,}144 & Section~\ref{sec:scale-companions} \\
    In-flight floor, at scale & 8 & 262{,}144 & Section~\ref{sec:scale-companions} \\
    Concurrency, at scale   & 8  & 262{,}144 & Section~\ref{sec:scale-companions} \\
    Clumpy-field, at scale  & 8  & 200{,}000 & Section~\ref{sec:scale-companions} \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Baseline validation}
By construction the ``instant'' regime is the $c\to\infty$ limit of (\ref{eq:gate}), and it
reproduces the plain perfect-information fold bit-for-bit. Against Nicholson and Forgan's
published model \citep{nicholson-forgan-2013} the event-driven fold reproduces both headline
findings: gravitational slingshots explore about two orders of magnitude faster than powered
flight, and their surprising result that aiming for the nearest star beats aiming for the largest
boost on total exploration time holds, while the maximum-boost policy reaches the higher peak
speed. The speed-up agrees to the same order of magnitude but not to the digit, as one should
expect: we measure a factor of about $166$ on a $400$-star field, against their factor of about
$100$ (a reduction from $\sim\!10^8$ to $\sim\!10^6$ years) on a $200{,}000$-star field, so with a
field some $500$ times smaller the exact multiple is not expected to match. A coarse fixed step had
quantized the boosted hops and undercut even this to about twenty-fold; resolving the step recovers
the full factor, which is the same correction that removes the spurious fill-time tax.
