\section{The coordination tax}

The cost was measured as a paired ensemble: 32 seeded galaxies of 300 stars each, every
one filled twice - once with perfect information and once with the light-speed gate - so
that the only difference between the two runs is the delay. Comparing the time to settle
the entire field, the median slowdown from light-speed lag is:

\begin{itemize}
  \item \textbf{powered nearest-neighbour flight: about 0\%} - essentially immune;
  \item \textbf{nearest-star slingshot flight: about 30\%} (interquartile range roughly
    $+20\%$ to $+38\%$);
  \item \textbf{maximum-boost slingshot flight: about 50\%} (interquartile range roughly
    $+46\%$ to $+54\%$).
\end{itemize}

Two facts stand out. The first is that the penalty is \emph{not} simply $\Lambda = v/c$.
Powered flight is nearly untouched even though it is the slowest regime, and the two fast
slingshot regimes differ from each other by nearly a factor of two. If $\Lambda$ alone set
the cost, the fastest probes would pay the most in a smooth progression; instead the cost
tracks something else. That something is hop locality. A probe flying to the
\emph{nearest} star that loses a race has lost almost nothing: the next-nearest unsettled
star is right beside it, so the wasted trip and the recovery are both short. A probe
chasing the \emph{biggest boost}, by contrast, deliberately reaches past its neighbours to
a distant star; when it loses that race, both the wasted trip and the re-targeting detour
are long. So $\Lambda \approx v/c$ sets the \emph{scale} of the lag, but the
\emph{non-locality of each hop decides whether it bites}. Fast-but-local flight escapes;
fast-and-far flight pays.

The second fact is what does \emph{not} happen: every case still fills the field to
100\%. Light-speed lag makes the swarm slower and more wasteful, but it never leaves a
permanent unsettled remainder. Re-targeting guarantees that a probe which loses a race
simply picks another star and eventually the field is closed. A permanent settled fraction
below one - a steady state in which some region stays forever empty - requires settled
sites to \emph{die}, adding a loss term to the dynamics; delay by itself supplies no such
term. Lag is a tax on speed, not a barrier to completion.
