% main.tex - the coordination-tax paper. Hand-written; gen-tex.mjs leaves it alone.
% Journal-manuscript format: single-column article with author-year (natbib) citations,
% suitable as an initial submission to an astrobiology journal (e.g. the International
% Journal of Astrobiology, the venue of Nicholson & Forgan 2013, which this paper extends).
% Deliberately uses only TeX Live packages - no bundled publisher class, no biber - so CI
% typesets it unchanged; the publisher's own house template is applied at submission time.
% Bibliography is generated: ../refs.bib comes from frontend/src/sources.ts.
% Figures are generated (gitignored vector PDFs) by the swarm module via
% `uv run --extra dev python -m experiments.paper_figures`.
\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[round,authoryear]{natbib}
\usepackage{url}
\bibliographystyle{plainnat}

\title{The Coordination Cost of Light-Speed Delay in a\\ Self-Replicating Probe Swarm Scales as $v/c$}
\author{Noah Hyd\'en\\[2pt]
  \normalsize Independent researcher\\
  \normalsize ORCID: 0009-0003-4523-0467}
\date{}

\begin{document}
\maketitle

\begin{abstract}
A swarm of self-replicating probes fills a stellar field star by star, each probe settling a star
and launching copies. The standard exploration-timescale model grants every probe perfect,
instantaneous knowledge of which stars are already settled, and names finite light-speed as its own
future work. We add that light-speed limit and measure its cost. The first finding is a caution: a
model advanced in a timestep coarser than the interval between probe launches makes many probes
decide from the same stale snapshot at once, and the resulting collisions appear to slow the fill by
tens of percent. That number is a discretization artifact. Resolved so that each probe decides at
its true arrival time, the tens-of-percent slowdown collapses -- to a few percent at directed-energy
speeds and to nothing at slingshot speeds -- and the gravitational-slingshot speed-up survives,
reproducing the source model to the same order of magnitude (about 166 times powered flight here,
against its factor of about 100 on a field 500 times larger). The dominant cost is
redundant travel: with essentially the same number of probes built, stale views send probes to stars
others have already claimed. We show, and derive from a short collision argument, that stale views
multiply the wasted journeys by 1 + v/c, so the coordination tax -- the extra wasted journeys as a
fraction of those a perfect-information swarm already flies -- equals, at the default replication
branching, the probe's speed in units of
light, v/c: negligible for a powered cruise, about a percent at slingshot speeds, and a fifth at the
directed-energy speeds beamed propulsion targets. The perfect-information baseline is itself wasteful
-- about four in five journeys are redundant even with an instantaneous map -- so this is a v/c
surcharge on an already-redundant process, not a claim that a fifth of all journeys are wasted. The
tax grows with the replication branching factor, which sets the coefficient of the v/c law --
without saturating up to sixteen offspring, where it reaches about a third -- does not grow with
system size over the range we test, holds under a
clumpy (non-uniform) field where if anything it softens, and is not a hard floor: a probe that
listens in flight, rather than only at its decision stars, recovers essentially all of it at a small
cost in fill time. Whether the wasted travel also costs energy is mission-dependent, since a slingshot
probe's kinetic energy is largely gravitationally sourced, so we denominate the tax in journeys. The
coordination cost a perfect-information model hides is dominated by fuel and effort and scales,
cleanly, as v/c.
\end{abstract}

\noindent\textbf{Keywords:} self-replication; interstellar exploration; light-speed delay;
multi-agent coordination; swarm dynamics

\bigskip

\input{body/introduction}
\input{body/the-light-delayed-belief-model}
\input{body/the-coordination-tax}
\input{body/discussion}
\input{body/conclusion}

\bibliography{../refs}

\end{document}
