\section{The closure model}

We treat the factory as a bill of materials and define its closure $C$ as the mass
fraction of its own parts it can manufacture on site. The complement, $1-C$, is the mass
fraction that must be imported as vitamins for every kilogram the factory builds. Mass
balance makes this exact rather than approximate: to add one kilogram of installed
capacity, the factory must be supplied with $1-C$ kilograms it cannot make. This is
material closure; von Neumann's construction shows it is inseparable from information
closure, since a machine that cannot copy the description of its chips has not truly closed
on them either \cite{von-neumann-burks-1966,freitas-merkle-2004}.

The literature does not place $C$ at one. The canonical seed-factory study estimated that
\mbox{90-96\%} mass closure is achievable and argued that chasing the last few percent to
full closure is not worthwhile \cite{nasa-cp-2255-1980}. More recent designs are more
conservative: a colony-scale factory takes a near-term-realistic closure of about 70\% with
chips, solar cells, and circuitry staying Earth-sourced
\cite{guided-self-replicating-factory-2021}, and an interstellar-probe design lands at a
similar 70\% replicated mass fraction with electronics imported \cite{borgue-hein-2020}.
The modern synthesis agrees on the principle: because the marginal part gets
disproportionately harder to make as closure rises, the sensible target is
high-but-incomplete closure, not one \cite{freitas-merkle-2004}.

Across these designs the conclusion is the same. Imported electronics are a permanent
design feature, not a temporary compromise on the way to full autonomy. The interesting
question is therefore not \emph{whether} a factory imports electronics but \emph{why} that
particular part is the one it cannot make, and what the resulting cap on closure costs. The
answer rests on an energy figure, so we state its basis up front: throughout, embodied
energy means the electrical energy to manufacture one kilogram of a finished, packaged part
on-site, and for chips specifically the finished packaged part rather than the bare silicon
die. That basis matters, because the same object's embodied energy differs by more than an
order of magnitude depending on where the boundary is drawn.
