\section{The electronics wall}

Chips are the part a lone factory cannot make, and the reason is twofold. The first
foundation is supply-chain depth. Semiconductors sit at the end of the deepest supply
chain on Earth: more than 400 process steps, materials purified to nine-nines purity,
and a global monopoly on the extreme-ultraviolet lithography that patterns the
smallest features \cite{csis-semiconductor-supply-chain}. No single factory,
however well closed on structure, reproduces that pipeline from local rock.

The second foundation is energy, and it is the one that turns a supply-chain argument
into a physical wall. The embodied energy of compute logic is on the order of
8,000~kWh/kg, against roughly 5~kWh/kg for smelted metal - a gap of about three orders
of magnitude \cite{williams-ayres-heller-2002,nagapurkar-das-2022}. A factory that
would make its own chips must therefore pour vastly more electricity into a kilogram
of logic than into a kilogram of the structure that houses it.

These two foundations combine into a single conditional. A factory escapes the
electronics wall only if it is both highly self-sufficient \emph{and} swimming in
power, and the coupling is sharp. For a highly closed lunar seed at 97\% closure,
letting it make its own chips locally cuts the time to reach a target scale from about
29 years to about 17 years, but only when it is fed on the order of 4~MW; at about
1~MW, making chips locally backfires, and the factory runs out of electricity before
it runs out of parts \cite{nasa-cp-2255-1980}. Power is not a free knob. The energy
cost of the chips is what decides whether local production helps or hurts, and only a
factory already swimming in power comes out ahead.
