\section{Discussion}
\label{sec:discussion}

\subsection{Limitations}
The load-bearing limitation is the numerator. The copy time of \eqref{eq:copy} is computed
from a lunar-regolith seed because no per-module mass and energy breakdown yet exists for a
concrete interstellar-probe design; the most developed such design reports its architecture
but not a closed bill of materials at the fidelity this calculation
needs~\citep{borgue-hein-2020}, which is why we stand on a lunar factory rather than its probe
numbers. We therefore do not present 582~days as a probe's copy time. The paper's defense
against this gap is the margin of Fig.~\ref{fig:margin}, and it is honest about being bounded:
for the powered galactic fill the conclusion holds for any copy time within about two orders of
magnitude of the proxy (the per-copy ratio $f$ survives four orders, but the cumulative cost is
the binding measure), which comfortably spans the range a heavier, less-closed, or more
power-starved probe could plausibly occupy; for the fastest slingshot fronts the margin is only
a few-fold, so there a substantially slower probe could make manufacturing a percent-level
effect. One omission bears on exactly that boundary: the copy mass $C\,m_\mathrm{seed}$ counts
the offspring's local structure but not the multi-megawatt power plant the seed scenario treats
as separate infrastructure, which can outmass the seed itself. Folding it in would raise the
copy time, plausibly by up to about two orders of magnitude; that stays inside the powered
margin but would push the fastest slingshot fronts past one percent, consistent with the stated
boundary. Closing the gap with a sourced probe bill of materials would replace the proxy with a
measurement, but on the present evidence it cannot change the verdict for the powered galactic
case.

The dwell is evaluated at 1~AU around a Sun-like star, matching the uniform field the
settlement-front model assumes but not the real spread of stellar luminosities and material
availability. Because the copy time is machinery-limited at 1~AU, stellar luminosity does not
enter it at all until the insolation deficit exceeds the roughly 190-fold machinery-to-power
margin, beyond about 13.7~AU-equivalent, so a fainter star does not lengthen the copy time
across essentially the whole target population. Material availability enters through the closure
ratio and the machinery rate, that is, the numerator, so its effect is bounded by the same
margin. We deliberately do not add a stellar-population sub-model: it would be a separate slice
with its own validation, not a parameter of this one, and the margin makes it unnecessary for
the present claim. Likewise we
inherit the front model's idealizations, chiefly a frozen stellar field, but our claim is a
ratio in which the dwell enters the same way whether or not those idealizations hold, so it is
insulated from them in a way an absolute fill time would not be.

Finally, the direct tax for max-boost slingshots is unresolved at the field size we can
afford (Fig.~\ref{fig:tax}); we report it as positive but within seed noise rather than as a
number. That honesty is the point of the ensemble. The powered tax is smaller still and is
knowable only analytically, through \eqref{eq:frac}, not by brute force.

\subsection{Implications}
The substantive implication is a division of labour between scales. Manufacturing cadence is
the binding constraint on a local, densely packed fleet, where it sets the doubling time
outright, and is a negligible tax on a sparse interstellar front, where propulsion policy and
raw transit time dominate the timeline by nearly five orders of magnitude for a powered cruise,
narrowing to about two for the fastest slingshots. A programme that worries about how fast its
seeds can build is asking the right question for the solar-system phase and, for a powered
galactic cruise, the wrong one for the galactic phase, where the same effort is better spent on
propulsion; only when the front is driven to slingshot or directed-energy speeds does the build
cadence begin to re-enter the galactic ledger. The
result also illustrates a modelling stance: by keeping the three scales as separate, honestly
sourced models joined at a clean seam, rather than one fused simulation, we can derive a shared
quantity in the place it is physically fixed and route it to the place that needs it, and read
off which constraint binds where. The seam is what makes the cross-scale statement possible;
fusing the scales would have buried it.
