\section{The cross-scale model}
\label{sec:model}

\subsection{One factory, three folds}
Each scale is a separate, deterministic model that we drive from a single shared factory
specification. The factory is the sourced lunar-regolith seed of the closure and
replication literature~\citep{nasa-cp-2255-1980,metzger-2013,chirikjian-2004-niac,moses-chirikjian-2020}:
a mass-closure figure $C$ (the fraction of its own mass it can remake locally) and a bill
of materials whose masses and per-part energies are individually cited. We use it as a
stand-in for a probe because a probe-specific bill of materials is not yet available in the
literature (see Section~\ref{sec:discussion}); the point of this paper is that the choice
does not matter to the conclusion.

The leverage that motivates the whole programme follows from mass balance alone: if a factory
must import $(1-C)$~kg of parts for every kilogram it builds, then installed mass per launched
mass is $1/(1-C)$. At the fixed, very large velocity change of an interstellar transfer the
rocket equation charges a fixed propellant multiple $e^{\Delta v / v_e}-1$ for every kilogram
delivered~\citep{tsiolkovsky-1903}, so cutting the delivered mass by the leverage factor
$1/(1-C)$ cuts the propellant bill by that same factor: a linear saving on a kilogram that is
exponentially expensive to move. Modern seed-factory studies place achievable closure in the
range of roughly 70 to 96 percent and agree that chasing full closure is not
worthwhile~\citep{guided-self-replicating-factory-2021,nasa-cp-2255-1980,freitas-merkle-2004},
so a high but sub-unity $C$ is the realistic regime.

\subsection{The one derived number}
The quantity every scale shares is the time to build one copy. A factory builds at a rate set
by the lesser of its machinery throughput and the power its collectors can supply, and the time
to lay down one offspring's worth of local structure is
\begin{equation}
t_\mathrm{copy} \;=\; \frac{C \, m_\mathrm{seed}}{\min\!\left(\dot{m}_\mathrm{mach},\,\dot{m}_\mathrm{power}\right)},
\label{eq:copy}
\end{equation}
where $m_\mathrm{seed}$ is the seed mass, $\dot{m}_\mathrm{mach}$ the machinery-limited build
rate, and $\dot{m}_\mathrm{power}$ the rate the available power supports at the per-kilogram
embodied energy of the bill of materials (a metals-dominated figure~\citep{ice-coefficients,gutowski-2009},
with integrated circuits the high-intensity minority~\citep{nagapurkar-das-2022}). We evaluate
\eqref{eq:copy} at 1~AU around a Sun-like star, consistent with the uniform field the
settlement-front model assumes, where the solar constant is
$1361\ \mathrm{W\,m^{-2}}$~\citep{kopp-lean-2011}. It matters which of the two terms binds. For
the default seed the machinery throughput does: it lays down about 20~kg per day, roughly 190
times below the rate the 1-AU array could power, so $t_\mathrm{copy}$ is fixed by machinery
throughput and $C\,m_\mathrm{seed}$ alone, and the power branch would take over only past about
13.7~AU, where the $1/d^{2}$ insolation has fallen to the machinery rate. The copy time is
therefore insensitive to insolation across essentially the whole target population, which
strengthens rather than weakens the cross-scale claim below. This is the same expression the
fleet model already uses for its doubling cadence; nothing is added here.

The settlement-front model needs this number as a per-star manufacturing dwell in years,
\begin{equation}
\tau \;=\; t_\mathrm{copy} / Y, \qquad Y = 365.25\ \mathrm{d},
\label{eq:dwell}
\end{equation}
where $Y$ is the Julian year ($1\ \mathrm{yr} = 3.15576\times10^{7}\ \mathrm{s}$). The
same Julian year defines the speed of light in the units the front uses for light travel
and stellar distances, so the dwell and the fill time it will be compared against are on
one clock. A silent mismatch here would rescale the central ratio by orders of magnitude
without any error elsewhere, so we hold it as a machine-checked invariant of the code.

\subsection{The finding, reduced to one ratio}
With the dwell derived rather than zeroed, the front is run to completion and yields a
fill time $T_{100}$, the time for the last star in the field to be settled. The cleanest
single statement of scale separation is the dimensionless ratio
\begin{equation}
f \;=\; \tau / T_{100},
\label{eq:frac}
\end{equation}
one manufacturing dwell as a fraction of the whole galactic fill. This is a \emph{per-copy}
measure, and it is important not to mistake it for the total. The front pays one dwell at every
settlement, so the manufacturing share of the whole timeline is not $f$ but the cumulative
slowdown of the fill, which is $f$ multiplied by the number of settlements on the critical path
from the seed to the last star. We report both: $f$ fixes the order of magnitude and is the
quantity the robustness margin is stated on, while the cumulative cost is measured directly,
front against front, in Section~\ref{sec:results}. If $f \ll 1$ and the critical path is not
astronomically long, both of which we show, the build cadence that governs the local fleet,
where it is the only clock that matters because transit between neighbours is days, is a
negligible tax at galactic scale, where transit is millennia. The rest of the paper measures
$f$, tests how much it can be trusted given that $t_\mathrm{copy}$ is a proxy, and measures the
dwell's cumulative cost on the front directly.
